---
title: 'Exercises Data Wrangling 2 - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  #unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

For the following exercises we will also data from [*Gapminder*](https://www.gapminder.org/); this time on life expectancy.

As per usual, we first need to read in the data. You can just copy, paste and run the following code in(to) your script.

```{r read gap_life}
library(readr)

gap_life <- read_csv("../data/gapminder/life_expectancy_years.csv")
```

Again, the data are currently in wide format.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Select only data for the 20th century, but this time use a helper function instead of specifying a range columns.
```

```{block, opts.label = "clues"}
The helper function you should use here is `starts_with()`. We also want to keep the country column.
```

```{r starts with, solution = TRUE}
library(dplyr)

gap_life %>% 
  select(country, starts_with("19"))
```

As you may have already noticed, the dataset some missing data points. Before we start analyzing the data we might want to know for how many countries we have complete data.

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Using the dataset in wide format, find out for how many countries we have complete data?
```

```{block, opts.label = "clues"}
To answer this question you should use the `drop_na()` function from `tidyr`.
```

```{r drop_na gather, solution = TRUE}
library(tidyr)

gap_life %>% 
  drop_na() %>% 
  nrow()
```

As in the previous set of data wrangling exercises, we now want to transform the data into long format.

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Transform the *gap_life* dataset into a sensible long format. Name the variable representing the values for life expectancy *lifeExp* and store the resulting dataframe in a name with the same object as before (*gap_life*).
```

```{block, opts.label = "clues"}
This is just a repetition from the Tidy Data exercises. What we want to do is to gather the columns with the years into a year variable.
```

```{r gap_life gather, solution = TRUE}
library(tidyr)

gap_life <- gap_life %>% 
  gather(-country, key = "year", value = "lifeExp")
```

Now let's apply some of the advanced filtering options we discussed in the *Data Wrangling - Part 2* session.

```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Create two new dataframes that include different subets of the *gap_life* data: 1. Data for all countries for 1990s (name this one *gap_life_1990s*), 2. Data for all years but only for Germany (name this one *gap_life_ger*). NB: There are different Germanies in the dataset: *West Germany*. *East Germany*, and *Germany*. Before we can create the first new dataset, we need to ensure that the *year* variable is of the right type. To achieve this you can just copy, paste, and execute the following lines of code.
```

```{r data for exercise file, include = F}
library(tidyr)
library(dplyr)

gap_life <- gap_life %>% 
  gather(-country, key = "year", value = "lifeExp")
```


```{r gap_life change var type}
gap_life <- gap_life %>% 
  mutate(country = as.factor(country),
         year = as.integer(year))
```


```{block, opts.label = "clues"}
You need to use a helper function from `dplyr` to create the first new dataframe and a specific matching operator to create the second one. 
```

```{r gap_life filter advanced, solution = TRUE}
gap_life_1990s <- gap_life %>% 
  filter(between(year, 1990, 1999))

gap_life_1990s

gap_life_ger <- gap_life %>% 
  filter(country %in% 
           c("Germany", "West Germany", "East Germany"))

gap_life_ger
```

For some comparisons (especially via plots), it might help to know which continent the country is located on. For this purpose, we will create a new *continent* variable. As it would be quite tedious to create this variable manually for all of the countries in the dataset, we will do this only for a subset in this exercise. Just run the following code in your local script to create this subset.

```{r gap_life countries subset}
gap_life_subset <- gap_life %>% 
  filter(country %in% 
           c("Netherlands", "Brazil", "China", "Algeria", "New Zealand"))
```

```{block, box.title = "5", box.body = list(fill = "white"), box.icon = "fa-star"}
Create a *cotinent* variable for the countries in *gap_life_subset*. The variable should be a factor and its values the following: *Africa*, *Americas*, *Asia*, *Europe*, *Oceania*.
```

```{block, opts.label = "clues"}
You should use the `case_when()` to create this new variable.
```

```{r gap_life case_when, solution = TRUE}
gap_life_subset %>% 
  mutate(continent = factor(case_when(
    country == "Algeria" ~ "Africa",
    country == "Brazil" ~ "Americas",
    country == "China" ~ "Asia",
    country == "Netherlands" ~ "Europe",
    country == "New Zealand" ~ "Oceania")
    ))
```