---
title: 'Exercises Relational Data - Solutions'
#title: 'Exercises Relational Data'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  #unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

We will, again, use data from [*Gapminder*](https://www.gapminder.org/) for the exercises on relational data.  

There is an `R` package named [`gapminder`](https://cran.r-project.org/web/packages/gapminder/) that contains a selection of the data from *Gapminder*.

Our aim for the following exercises is to use the individual data files from *Gapminder* that we have used in some of the previous exercises to create combined datasets that resemble the one contained in the `gapminder` package.

Before we can begin with these exercises, we need to load the datasets. To do this, you can simply use the following code (that repeats some parts of the previous exercises).

```{r load gapminder datasets, message = F}
library(readr)
library(dplyr)

gap_cont <- read_csv("../data/gapminder/countries_continent.csv")

gap_life <- read_csv("../data/gapminder/life_expectancy_years.csv")

gap_pop <- read_csv("../data/gapminder/population_total.csv") %>% 
  rename(country = "Total population")

gap_gdp <- read_csv("../data/gapminder/gdppercapita_us_inflation_adjusted.csv")
```

First of all, let's have a look at the *Gapminder* data that is included in the `gapminder` package.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the `gapminder` package and have a look at the dataset it contains. The dataset we want is simply called *gapminder*.
```

```{block, opts.label = "clues"}
In order to print the dataset, you can just type its name and run that line.
```

```{r gapminder pkg, solution = TRUE}
library(gapminder)

gapminder
```

In the following, we will use different joins to create datasets that contain the same set of variables. We will create two versions of the combined dataset.  

Before we do this, however, we want to explore the overlap and discrepancies between the individual datasets. This is somewhat easier to do with the datasets in wide format (as each country name only appears in one row in those).

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Using a filtering join, determine which countries are in *gap_pop* but not in *gap_life* and vice versa. Do the same for *gap_gdp* and *gap_life*.
```

```{block, opts.label = "clues"}
You should use `anti_join()` for this task. To just get the country names, you can select the country variable.
```

```{r anti_join, solution = TRUE}
library(dplyr)

gap_pop %>% 
  anti_join(gap_life, by = "country") %>% 
  select(country)
  
gap_life %>% 
  anti_join(gap_pop, by = "country") %>% 
  select(country)

gap_gdp %>% 
  anti_join(gap_life, by = "country") %>% 
  select(country)
  
gap_life %>% 
  anti_join(gap_gdp, by = "country") %>% 
  select(country)
```

For the following series of joins, we want the data in the long format again. Just copy, paste, and run the code below to transform the datasets accordingly.

```{r wrangle gapminder datasets}
library(tidyr)

gap_life <- gap_life %>% 
  gather(-country, key = "year", value = "lifeExp") %>% 
  mutate(year = as.integer(year))

gap_pop <- gap_pop %>% 
  gather(-country, key = "year", value = "pop") %>% 
  mutate(year = as.integer(year),
         pop = as.integer(pop))

gap_gdp <- gap_gdp %>% 
  gather(-country, key = "year", value = "gdpPercap") %>% 
  mutate(year = as.integer(year))

```

As stated above, we want to create two different versions of the combined datasets: One without missing data and one with as many observations (rows) as possible.

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Create the two versions of the combined dataset described above using one type of mutating join for each one.  

For the one without missing data you should start with the dataset with the largest number of countries in it (*gap_pop*), and then join the other datasets in descending order of the number of countries they contain (*gap_life*, *gap_gdp*, *gap_cont*). This datasets also requires two additional (pipe) steps to ensure that a) it contains no NAs and b) its variables are in the same order as in the `gapminder` package dataset.  

For the dataset with the maximum amount of observations (rows), it does not matter which dataset you start from. Name the resulting datasets *gap_nomiss* and *gap_full*.
```

```{block, opts.label = "clues"}
The types of mutating joins that you should use are `inner_join()` and `full_join`. The variables you should join on are *country* and *year*.  

Remember that the *gap_nomiss* dataset requires two additional steps.
```

```{r join datasets, solution = TRUE}
gap_full <- gap_cont %>% 
  full_join(gap_life, by = "country") %>% 
  full_join(gap_pop, by = c("country", "year")) %>% 
  full_join(gap_gdp, by = c("country", "year"))

gap_nomiss <- gap_pop %>% 
  inner_join(gap_life, by = c("country", "year")) %>% 
  inner_join(gap_gdp, by = c("country", "year")) %>%
  inner_join(gap_cont, by = "country") %>% 
  drop_na() %>% 
  select(country, continent, year, lifeExp, pop, gdpPercap)
```

Now that we have created the two combined datasets let's see how much they differ from each other.

```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Using a set operation, determine how many observations from the *gap_full* dataset are missing from *gap_nomiss*.
```

```{block, opts.label = "clues"}
You can use `nrow()` to get the number of resulting observations for the set operation.
```

```{r set ops, solution = TRUE}
gap_full %>% 
  setdiff(gap_nomiss) %>% 
  nrow()
```

Another way to explore the differences between the datasets could be to look at some of the summary statistics for them. How to create summary statistics with the `tidyverse` will be the topic of the next session.

As a final exercise, let's join another variable (*average number of children per woman*) to the dataset from the `gapminder` package. To load and prepare the data, copy and run the following code.

```{r gap_fert prep, message = F}
library(gapminder)

gapminder <- gapminder %>% 
  mutate(country = as.character(country))

gap_fert <- read_csv("../data/gapminder/children_per_woman_total_fertility.csv") %>% 
  gather(-country, key = "year", value = "fert") %>% 
  mutate(year = as.integer(year))
```

```{block, box.title = "5", box.body = list(fill = "white"), box.icon = "fa-star"}
Join the *fert* variable from *gap_fert* to the *gapminder* dataset.
```

```{block, opts.label = "clues"}
The *gapminder* dataset should be the reference/master table for this join.
```

```{r left_join, solution = TRUE}
gapminder %>% 
  left_join(gap_fert, by = c("country", "year"))
```
