---
title: 'Exercises Tidy Data'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html: default
  unilur::tutorial_html_solution: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

You can write some text appearing in both files using the standard [markup language](http://rmarkdown.rstudio.com).

Use the `r` chunk to display some code in both files.

```{r}
head(cars)
```

Use the `r` chunk with `solution = TRUE` to display and evaluate the chunk only in the solution file.

```{r, solution = TRUE}
# Only visible and evaluated in the solution file...
plot(cars)
```

You can also add some markdown formatted text with the asis engine together with `solution = TRUE`

```{block, solution = TRUE}
This is the text explaining how to obtain the solution and written in **Rmarkdown**
```

Use the knit dropdown menu on the knit button to generate the desired output:

- Using `unilur::tutorial_pdf_solution` or `unilur::tutorial_html_solution` will show the solution boxes.
- Using `unilur::tutorial_pdf` or `unilur::tutorial_html` will hide them from the output.

```{block, box.title = "Tip", box.body = list(fill = "lightblue")}
You can also include custom boxes to highlight some content. Read the [article](http://koncina.github.io/unilur/articles/custom-boxes.html) to learn more about how to customize the appearance.
```

- What is the mean speed reported in the `cars` dataset?

```{r, solution = TRUE}
mean(cars$speed)
```

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the **tidyr** library
```

```{block, opts.label = "clues"}
Use `library()`.
```