---
title: 'Exercises Importing Data - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  #unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

After succesfully accomplishing the exercises on tidy data and listening to lengthy lectures on data formats as well as specifics of importing them, it's now your turn to get used to importing data in the `tidyverse`.

We prepared some datasets which you can use to play with some of the functions, e.g., from `readr`. You can find them in the `../data` folder. However, importing data often implies firing up only one command and that's it. For this reason, in these exercises, we prepared some special tasks you can work on.

