---
title: 'Exercises Importing Data - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  #unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

After succesfully accomplishing the exercises on tidy data and listening to lengthy lectures on data formats as well as specifics of importing them, it's now your turn to get used to importing data in the `tidyverse`.

We prepared some datasets which you can use to play with some of the functions, e.g., from `readr`. You can find them in the `../data` folder. However, importing data often implies firing up only one command and that's it. For this reason, in these exercises, we prepared some special tasks you can work on.



First, copy the following code into a new R script and run it to load/generate the datasets we will use in these exercises. For the moment, you do not have to understand all all of the following code (but you should be able to at the end of this workshop). Please note that in order for the code to run, your working directory should be the folder with the workshop materials (use `getwd` to print your current working directory if you are unsure and `setwd` to change it if necessary).  

```{r load data, message=FALSE}
library(tidyverse)

gap_life <- read_csv("../data/gapminder/life_expectancy_years.csv")
titanic <- read_csv("../data/titanic/titanic.csv")

crime <- tibble(country = rep(c("Germany", "Brazil", "Norway"), 2),
                  crime = c(rep("murders", 3), rep("intentional homicide rate", 3)),
                  year = 2010,
                  value = c(690, 40974, 29, 0.84, 27, 0.68))
```

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Have a look at the *gap_life* dataset. What do you notice?
```

```{block, solution = TRUE}
The dataset is in *wide format* with each year as a column/variable. To increase human-readability and to facilitate further data wrangling and exploration steps, it makes sense to transform it into *long format*.
```

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Transform the *gap_life* dataset into a sensible long format.
```

```{block, opts.label = "clues"}
You should `gather` the years into one column/variable. If you are unsure about the arguments of a function, you can always consult the help files by typing (and running) a ? directly followed by the function name (e.g., `?glimpse`). NB: This only works if you have previously loaded the package that includes the function.

Be aware that there is one column in the dataset that you do not want to gather. You can specify this by adding -variablename as the second argument (or first if you use pipes) to the `gather()` function.
```

```{r gap long, solution = TRUE}
gap_life_long <- gap_life %>% 
  gather(-country, key = "year", value = "life_exp")

gap_life_long
```

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Have a look at the *crime* dataset. What do you notice?
```

```{block, solution = TRUE}
Observations are spread across two rows. The intepretation of the *value* column depends on the value of *crime*.
```

```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Tidy the *crime* dataset, so that there is only one observation/row for each country.
```

```{block, opts.label = "clues"}
You should `spread` the *crime* variable.
```

```{r crime spread, solution = TRUE}
crime_spread <- crime %>% 
  spread(key = "crime", value = "value")

crime_spread
```

```{block, box.title = "5", box.body = list(fill = "white"), box.icon = "fa-star"}
Split the *Name* variable in the *titanic* dataset into two variables: one that contains only the last name, and one that contains the first name(s) plus title (Mr., Mrs., Dr., etc.).
```

```{block, opts.label = "clues"}
When looking at the *Name* variable in the *titanic* dataset you should notice that the last name is separated from the rest of the name by a comma plus a space.
```

```{r titanic names, solution = TRUE}
titanic <- titanic %>% 
  separate(Name, c("last_name", "first_name"), sep =", ")

titanic
```