---
title: 'Exercises Importing Data - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  #unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

After succesfully accomplishing the exercises on tidy data and listening to lengthy lectures on data formats as well as specifics of importing them, it's now your turn to get used to importing data in the `tidyverse`.

We prepared some datasets which you can use to play with some of the functions, e.g., from `readr`. You can find them in the `../data` folder. However, importing data often implies firing up only one command and that's it. For this reason, in these exercises, we prepared some special tasks you can work on.


```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the titanic data.
```

```{block, opts.label = "clues"}
The file format is CSV, accordingly you need the `readr` library and the function `read_...`
```

```{r load_titanic, solution = TRUE}
library(readr)

titanic <-
  read_csv("../data/titanic/titanic.csv")
```



```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Convert the variable `Sex` to a factor.
```

```{block, opts.label = "clues"}
You can do that while importing the data or after loading them.
```

```{r convert_sex, solution = TRUE}
titanic <-
  read_csv(
    "../data/titanic/titanic.csv",
    col_types = cols(
      Sex = col_factor()
    )
  )
```


```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the gapminder GDP data.
```

```{block, opts.label = "clues"}
I have no clue.
```

```{r load_gapminder_gdp, solution = TRUE}
library(readxl)

gapminder_GDP <-
  read_excel("../data/gapminder/GDPpercapitaconstant2000US.xlsx")
```



```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the unicorn sales data file. As we are not interested in the `total_turnover` variable only read in the cell range `A1:C43`
```

```{block, opts.label = "clues"}
You can define ranges with the option `range = range_definition`.
```

```{r load_unicorn_sales, solution = TRUE}
library(readxl)

unicorn_sales <-
  read_excel(
    "../data/unicorns/sales.xlsx",
    range = "A1:C43"
      )
```
