---
title: 'Exercises Summary Statistics - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  unilur::tutorial_html: default
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))

library(readxl)
library(dplyr)

gapminder_example <- 
  readxl::read_excel(
    path = "../data/gapminder/GDPpercapitaconstant2000US.xlsx",
    sheet = "Data"
    )
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

Imagine we're some fancy data scientists exploring - once again - the gapminder data. We're particularly interested in the development of the GDP across time *and* across countries. Some `R`-fanatics from GESIS recommended using this `tidyverse` thing in order to complete our tasks. At the same time, they also hesitate to load all of its `R`-packages at once.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load all packages from the `tidyverse` for importing Excel data and for data wrangling.
```

```{block, opts.label = "clues"}
You can find them in the slides from A3 and A5.
```

```{r load_packages_solution, solution = TRUE, message = FALSE}
library(readxl)
library(dplyr)
library(tidyr)
```

Ok, that wasn't too hard. But data science is about data, so we have to load in the data.

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Import the GDP gapminder data. Make sure to only import the excel sheet named "Data".
```

```{block, opts.label = "clues"}
Individual sheets can be chosen by applying the option `sheet = "name_of_your_sheet"`
```

```{r load_gapminder_solution, solution = TRUE, message = FALSE}
gapminder_example <- 
  readxl::read_excel(
    path = "../data/gapminder/GDPpercapitaconstant2000US.xlsx",
    sheet = "Data"
    )
```

Have the data been succesfully imported? They should comprise a tibble of 275 x 53. Furthermore, the income per person for Algeria of the years 1960, 1961, and 1962 should be 1280, 1085, and 856.

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Proof that the income per person for Algeria of the years 1960, 1961, and 1962 are 1280, 1085, and 856
```

```{block, opts.label = "clues"}
Algeria is in the 5th row of the dataset and the relevant variables are in the first four columns. You can subset datasets also by selecting rows by number with `select()` and by filtering by number with `slice()`.
```

```{r find_algeria_solution, solution = TRUE, message = FALSE}
gapminder_example %>% 
  select(1:4) %>% 
  slice(n = 5)
```




```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Task: What is the mean GDP of 1960-1969 over all countries?
```

```{block, opts.label = "clues"}
...
```

```{r mean_all_solution, solution = TRUE}
gapminder_example %>% 
  mutate(country = `Income per person (fixed 2000 US$)`) %>% 
  select(country, starts_with("196")) %>% 
  gather(-1, key = "year", value = "GDP") %>%
  filter(!is.na(GDP)) %>% 
  arrange(year, GDP) %>%
  group_by(year) %>% 
  summarise(`GDP_over_all_countries` = mean(GDP))
```

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Transform the *gap_life* dataset into a sensible long format.
```
