---
title: 'Exercises Terminology - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
#unilur::tutorial_html: default
---
  
  ```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

This is the very firt exercise of this course. It's not too hard, its purpose is rather to get used to our exercise format - this HTML file, for instance - and more importantly, to get used to the `tidyverse` world.

Consequently, this exercise is really short and just let you play around with some pipes and tibbles. It's a mini-exercise!

First things first: To work with the 'tidyverse' we have to have access to its packages.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the `tidyverse` library.
```

```{block, opts.label = "clues"}
If the `tidyverse` library has not been installed yet, you can install it with the command `install.packages("tidyverse")`.
```

```{r load_tidyverse_solution, solution = TRUE, message = FALSE}
if (!require(easypackages)) install.packages("easypackages")
library(easypackages)
```