---
title: 'Exercises Terminology - Solutions'
author: 'Johannes Breuer, Stefan Jünger, Thomas Ebel'
date: 'Data Wrangling & Exploration with the Tidyverse in R, May 15-16, 2019'
output:
  unilur::tutorial_html_solution: default
  unilur::tutorial_html: default
---
  
  ```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(clues = list(box.title = "Clues",
                                      box.body = list(fill = "#fff9dc", colour = "black"),
                                      box.header = list(fill = "#ffec8b", colour = "black"),
                                      box.icon = "fa-search",
                                      box.collapse = TRUE))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

This is the very firt exercise of this course. It's not too hard, its purpose is rather to get used to our exercise format - this HTML file, for instance - and more importantly, to get used to the `tidyverse` world.

Consequently, this exercise is really short and just let you play around with some pipes and tibbles. It's a mini-exercise!

First things first: To work with the 'tidyverse' we have to have access to its packages.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the `tidyverse` library.
```

```{block, opts.label = "clues"}
If the `tidyverse` library has not been installed yet, you can install it with the command `install.packages("tidyverse")`.
```

```{r load_tidyverse_solution, solution = TRUE, message = FALSE}
if (!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```

```{r to_pipe, message=FALSE}
mean(sqrt(as.numeric(read.csv2("../data/titanic/titanic.csv", sep = ",")$Fare)))
```



```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Create a nice pipe from this nested `base-R` command.
```

```{block, opts.label = "clues"}
You can call individual columns of a piped object with `.$col_name`.
```

```{r to_pipe_solution, , solution = TRUE}
read.csv2("../data/titanic/titanic.csv", sep = ",") %>% 
  .$Fare %>% 
  as.numeric() %>% 
  sqrt() %>% 
  mean()
```


```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the `titanic` dataset and convert it immediatly to a `tibble`. 
```

```{block, opts.label = "clues"}
`base-R`'s `read.csv2()` is your friend. Also, you may want to do it all at once in one pipe.
```

```{r titanic_to_tibble_solution, , solution = TRUE}
titanic_tibble <-
  read.csv2("../data/titanic/titanic.csv", sep = ",") %>% 
  as_tibble()
```

```{r to_tribble, message = FALSE}
x <- 1:10
y <- 100:91
z <- c(rep("bla", 5), rep("blubb", 5))

binded <- cbind(x, y, z)

binded_df <- as.data.frame(binded)
```


```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Use the `tribble()`-function to directly rebuild the data frame as a tibble
```

```{block, opts.label = "clues"}
Remember defining new columns with a preceding `~`. 
```

```{r to_tribble_solution, message = FALSE}

binded_tribble <-
  tribble(
    ~x,  ~y,   ~z, 
     1,   100,  "bla",
     2,    99,  "bla",
     3,    98,  "bla",
     4,    97,  "bla",
     5,    96,  "bla",
     6,    95,  "blubb",
     7,    94,  "blubb",
     8,    93,  "blubb",
     9,    92,  "blubb",
     10,   91,  "blubb"
  )
```



