---
title: "Data Wrangling & Exploration with the Tidyverse in R"
subtitle: "Outlook"
author: "Johannes Breuer<br />Stefan Jünger<br />Thomas Ebel"
date: "2019-05-16"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "../workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include = FALSE}
if (!require(easypackages)) install.packages("easypackages")
library(easypackages)

packages(
  "knitr", "rmarkdown", "magrittr", "readr", "ggplot2", "hadley/emo",
  prompt = FALSE
)

options(htmltools.dir.version = FALSE)

opts_chunk$set(echo = FALSE, fig.align = "center")
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", " & ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$title)`</span></div>
  </div>
  
---

# Course recap

```{r recap 1, echo = F}
recap_1 <- data.frame(
  "Session" = c(
    "Introduction", 
    "Tidyverse terminology: tibbles & pipes", 
    "Tidy data", 
    "Importing data", 
    "Data wrangling: Part 1"
    ),
  "Example content" = c(
    "What is the tidyverse and what is it good for?", 
    "%>% operator & tibbles", 
    "Gathering and spreading from wide to long and back, separating & uniting values", 
    "Importing csv, Excel, SPSS, and Stata files", 
    "Renaming, filtering, selecting, new variables, recoding, missing values, arranging"
  )
  , stringsAsFactors = FALSE
  , check.names = FALSE
)

kable(
  recap_1
  , format = "html"
  , align = "ll"
  , escape = FALSE
)
```

---

# Course recap

```{r recap 2, echo = F}
recap_2 <- data.frame(
  "Session" = c(
    "Data wrangling: Part 2", 
    "Relational data", 
    "Data exploration I: Summary statistics", 
    "Data exploration II: ggplot2 basics", 
    "Outlook"
    ),
  "Example content" = c(
    "Advanced selecting, filtering, creation and transformation of variables, strings and factors", 
    "Mutating joins, filtering joins, set operations", 
    "Grouping and summarizing data", 
    "Layers: Data, aesthetics, and geoms", 
    "What we did and what we did not cover in this workshop"
  )
  , stringsAsFactors = FALSE
  , check.names = FALSE
)

kable(
  recap_2
  , format = "html"
  , align = "ll"
  , escape = FALSE
)
```

---

# Where to go from here?

Some topics that we did not cover or only brielfy touched upon that you might want to explore next/further:

- working with strings (with [`stringr`](https://stringr.tidyverse.org/index.html))

- working with factors (with [`forcats`](https://forcats.tidyverse.org/))

- working with times and dates (with [`lubridate`](https://lubridate.tidyverse.org/))

- advanced plotting with [`ggplot2`](https://ggplot2.tidyverse.org/)

- tidy modeling (using the [`broom`](https://github.com/tidymodels/broom) package)

- reporting & publishing (with [`RMarkdown`](https://rmarkdown.rstudio.com) and [`knitr`](http://yihui.name/knitr))

- functional programming (with the [`purrr` package](https://purrr.tidyverse.org/))

---

# Acknowledgements `r ji("heart")`

All slides were created with the `R` package [`xaringan`](https://github.com/yihui/xaringan) which builds on [`remark.js`](https://remarkjs.com), [`knitr`](http://yihui.name/knitr), and [`R Markdown`](https://rmarkdown.rstudio.com). The exercises were created with the [`unilur` package](https://github.com/koncina/unilur).  

The `workshop.css` file we used for the layout of the slides includes elements from `CSS` files for `xaringan` presentations by [Frederik Aust](https://github.com/crsh) and [David Zimmer](https://github.com/DavZim).  

We thank Anne Carome from *RStudio* for sending us some nice hex stickers for this workshop, Claudia O'Donovan-Bellante and her team for taking good care of the organization of this workshop, and all of you for participating!



