---
title: "Data transformation"
subtitle: "Verbs that came, saw, and conquered?"
author: "Johannes Breuer<br />Thomas Ebel<br />Stefan Müller"
date: "2018-05-15"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "../footer-header.css", "../mycss.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include=FALSE}
if (!require(knitr)) install.packages("knitr")
if (!require(rmarkdown)) install.packages("rmarkdown")
if (!require(tidyverse)) install.packages("tidyverse")
library("knitr")
library("rmarkdown")
library("tidyverse")
options(htmltools.dir.version = FALSE)

opts_chunk$set(echo = FALSE, fig.align = "center")
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", " & ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$title)`</span></div>
</div>

---

```{r, include=FALSE, warning=FALSE, error=FALSE}
# devtools::install_github("DavZim/colorTable")
library(colorTable)

# define some colors for later
green <- "#a6d96a"
blue  <- "#74add1"
yellow <- "#f4e842"
```
class: center, middle

<h1> The tidyverse verbs</h1>
<h2> Or: How you never knew you couldn't live without them.</h2>

---

<h2>Verbs</h2>
The tidyverse comes with a range of new functions, called verbs, that make data analysis easier.  

These are especially:

* filter (rows)
* select (columns)
* arrange (the order of rows)
* mutate (i.e. create new columns)
* group_by/summarize (create summary/aggregate values)


.bottom[
There are a handyful of variations, like  
* mutate_if (create new variable dependend on some condition),
* rename (an existing variable and select all),  
and other functions, too.  
We'll get to some but not even close to all of them.
]

---

class: center, middle

<h2> Thanks!</h2>

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
