---
title: "Data transformation"
subtitle: "Verbs that came, saw, and conquered?"
author: "Johannes Breuer<br />Thomas Ebel<br />Stefan Müller"
date: "2018-05-16"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "../footer-header.css", "../mycss.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include=FALSE}
if (!require(knitr)) install.packages("knitr")
if (!require(rmarkdown)) install.packages("rmarkdown")
if (!require(tidyverse)) install.packages("tidyverse")
library("knitr")
library("rmarkdown")
library("tidyverse")
options(htmltools.dir.version = FALSE)

opts_chunk$set(echo = FALSE, fig.align = "center")#
set.seed(123)
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", " & ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$title)`</span></div>
</div>

---

```{r, include=FALSE, warning=FALSE, error=FALSE}
# devtools::install_github("DavZim/colorTable")
library(colorTable)

# define some colors for later
green <- "#a6d96a"
blue  <- "#74add1"
yellow <- "#f4e842"
```
class: center, middle

<h1> The tidyverse verbs</h1>
<h2> Or: How you never knew you couldn't live without them.</h2>

---

<h2>Verbs</h2>
The tidyverse comes with a range of new functions, called verbs, that make data cleaning (and analyzing) easier.  
These verbse are provided with the package dplyr.  

The most important verbs are:

* filter (rows)
* arrange (the order of rows)
* select (columns)
* mutate (create new columns)
* group_by/summarize (create summary/aggregate values)

--

.bottom[
There are a handful of variations, like  
* mutate_if (create new variable dependent on some condition),
* rename (an existing variable and select all),  
and other functions, too.  
]

---

<h2>The flow of things</h2>
All verbs follow the same logic:

Input is a data frame/tibble - do something with it - output is a new data frame/tibble.

--

E.g. the verb "filter"

```{r echo=TRUE, results='hide'}
table1 %>%
  filter(cases < 800)
```

--

1. Input is a tibble, table1
1. Filter this tibble for all rows with "cases" smaller than 800
1. Output is a new tibble

.bottom[
Note, that we are using Non-Standard-Evaluation (NSE) here. Instead of "table1$cases" or "table1[, "cases"] we just say 'cases' (without quotes!). More on that in A4.
]

---

<h2>Verb 'filter'</h2>
We already saw 'filter' in action. Now, let's explore it in depth.  

'filter' filters rows dependent on one or more conditions.  

--

You can use the usual operators for 
* comparisons: <, <=, ==, != (not equal), >=, >
   * e.g. 'filter(year != 1999)' and
* logical operators: &, |, ! (not), xor (either or, not both)
  * e.g. 'filter(year == 1999, cases < 3000)'

.bottom[
Note, that multiple conditions in filter are automatically added as '&' ('and').
]

---

<h3>Some examples for filter</h3>

<h4> Not equal </h4>

```{r echo=TRUE}
table1 %>%
  filter(year != 1999)
```

Output is a tibble with all rows where year is not equal to 1999.

---

<h4>Multiple conditions</h4>
```{r echo=TRUE}
table1 %>%
  filter(year == 1999, cases < 40000)
```

Output is a tibble with all rows where year is equal to 1999 and cases is smaller than 40.000.

--

An equivalent - but less canonic way - to express the same thing is 

```{r echo=TRUE}
table1 %>%
  filter(year == 1999 & cases < 40000)
```

---

<h4> Or </h4>
```{r echo=TRUE}
table1 %>%
  filter(year == 1999 | country == "Brazil")
```

Output is a tibble with all rows where either country is Brazil or year is 1999. If both are true that's okay, too. See the seconds row in the output.

---

<h4> Xor </h4>
```{r echo=TRUE}
table1 %>%
  filter(xor(year == 1999, country == "Brazil"))
```

Output very similar to the previous example but one row was dropped: The one that was true for both conditiones (year==1999, country=="Brazil"). That's the difference between '|' and 'xor'.

--

.bottom[
.highlight[Attention:] Never use base R's additional comparison operators '&&' and '||'. They only compare the first element of vectors and do not work with  tidyverse pipes!
]
---


<h2>Verb 'arrange'</h2>

'arrange' changes the order of rows in the data set.  

```{r echo=TRUE}
table1 %>%
  arrange(desc(population))
```

Arrange cases by their population value, in a decreasing order, from big to small.  

---

```{r echo=TRUE}
table1 %>%
  arrange(year, cases)
```

Arrange cases by year, then cases.

.bottom[
Note, that missings are always sorted at the end!
]

---

<h2>Verb 'select'</h2>

We can use 'select' to select the columns/variables that we want to look at.

--

Often we do not want to deal with all variables in a data set because there could be easily hundreds of variables (too many to be printed nicely in the console).

```{r echo=TRUE}
table1 %>%
  select(country, cases)
```

---

There are nice helper functions/operators for 'select'.

* colon --> 'select(table1, year.highlight[:]cases)' --> selects all variables **in the range** from year to cases

--

* minus --> 'select(table1, .highlight[-(]year:cases.highlight[)]' --> select all variables **besides** the mentioned ones/range.

--

* 'starts_with'/'ends_with' --> 'select(table1, .highlight[starts_with('w1')]) --> select all variables that start/end with string.

--

```{r echo=FALSE}
t_startswith <- tibble(
  'id' = 1:10,
  'w1_x' = sample(x = 1:10, replace = T),
  'w1_y' = sample(x = 1:10, replace = T),
  'w1_z' = sample(x = 1:10, replace = T),
  'w2_x' = sample(x = 1:10, replace = T),
  'w2_y' = sample(x = 1:10, replace = T),
  'w2_z' = sample(x = 1:10, replace = T),
  'w3_x' = sample(x = 1:10, replace = T),
  'w3_y' = sample(x = 1:10, replace = T),
  'w3_z' = sample(x = 1:10, replace = T)
)
glimpse(t_startswith)

```

---

* 'contains' --> 'select(table1, .highlight[contains('u')])' --> select all variables that contain the string.

--

* 'matches' --> 'select(t_startswith, .highlight[matches("^w\\\\\d_z")])' --> select all variables that match a regular expression.
```{r echo=TRUE}
select(t_startswith, matches("^w\\d_z"))
```

--

* everything() --> 'select(table1, cases, .highlight[everything()])' --> used to rearrange the order of variables (here, put "cases" in front).

---

<h2>Verb 'mutate'</h2>

With 'mutate' we create new variables based on existing ones.  

--

Creating the rate variable  

```{r echo=TRUE}
table1 %>%
  mutate(
    rate_in_perc = (cases / population) * 100
  )
```
---

You can already refer to newly created variables in 'mutate'

```{r echo=TRUE}
table1 %>%
  mutate(
    rate = cases / population,
    rate_in_perc = rate * 100 #<<
  )
```

---

Again, there are many useful helper functions/operators.  

* Of course, all arithmetic operators: +, -, /, *, ^, %% (modulo/remainder), %/% (integer devision)
  * they work, as you would expect in R, vectorized, i.e. applied to every element in the vector.

--

* aggregate functions: sum, mean, max, min, ...

--

* logs: log(), log2(), log10()

--

* offsets: lead(), lag()

--

* cumulative aggregates: cumsum, cumprod, cummin, cummax, cummean

--

* logical comparisons: <=, <, ==, !=, >, >=

--

* ranks: min_rank, ntile(), ...

--

That's a lot.  
But don't worry. We'll now go through some examples.

---

<h4> Arithmetic operators example</h4>

```{r echo=TRUE}
table1 %>%
  mutate(
    population_in_thousands = population / 1000
  )
```

--

.bottom[
Please note, that it is a deliberate choice of tibbles to not show 'insignificant' values behind the decimal point. https://github.com/r-lib/pillar/issues/40. Change with 'options(pillar.sigfig = 5)'.
]

---

<h4> Aggregate functions example</h4>

```{r echo=TRUE}
table1 %>%
  mutate(
    prop_cases = cases / sum(cases),
    max_cases = max(cases),
    mean_population = mean(population)
  ) %>%
  select(-country, -year)
```
--

  * 745 cases in Afghanistan are 0.136% of all cases (547660).
  * The most cases were reported in China 2000.
  * Mean population over all observations is 490072924.

.bottom[
Aggregate functions are not too useful on individual data. We'll get to grouping later.
]

---


<h4>offset example</h4>
```{r echo=TRUE}
offset_example <- tibble(
  hour = 8:17,
  productivity = c(80, 80, 120, 120, 60, 
                   70, 100, 100, 120, 80)) # index 100 is 'normal'
```

--

```{r echo=TRUE}
offset_example %>%
  mutate(change_in_prod = productivity - lag(productivity))
```

---

<h4> Cumulative agggregates example</h4>
```{r echo=TRUE}
table1 %>%
  mutate(
    cum_cases = cumsum(cases),
    prop_cases_perc = (cum_cases / sum(cases)) * 100 ,
    cum_mean_cases = cummean(cases)
  ) %>%
  select(contains("cases"))
```

---
<h4> Logical comparisons example</h4>

```{r echo=TRUE}
offset_example %>%
  mutate(
    unproductive = productivity <= 70,
    notable_deviations = 
      (abs(productivity - mean(productivity))) >= sd(productivity)
  ) %>%
  select(contains("product"), notable_deviations)
```

---

<h4> Rank example</h4>

```{r echo=TRUE}
table1 %>%
  mutate(
    ranks = min_rank(cases),
    ranks_rev = min_rank(desc(cases))
  )
```

---

<h2>Grouping and summarizing</h2>

We can group data by columns and then compute summary statistics/values.  

--
```{r echo=TRUE}
table1 %>% 
  group_by(year) %>%
  summarise(
    nr_cases = n(), min_cases = min(cases), max_cases = max(cases),
    mean_cases = mean(cases), sum_cases = sum(cases)
  )
```
--
Note that 
* there are only 2 observations (because year = [1999, 2000]).
* only the grouping variable and the newly created variables remain.

---

<h4>More on summarize functions</h4>

Ungrouping: 
* ungroup()

We can group by several variables:
* group_by(var1, var2, var3)


Useful functions:
* averages, quantiles: mean(), median(), quantile(x, 0.75)
* counts: n(), sum(!is.na()) (non-NA count), n_distinct() (unique values)
* measures of variations: 
  * sd() (standard deviation)
  * IQR() (interquatile range, difference between third quantile and first quantile), 
  * mad() (median absolute deviation, median of absolute deviations from median)
* position: first(), nth(x, 10), last()  


More on summarize functions in B3 tommorrow!


---

<h2> Your turn! Excercises :) </h2>

```{r}
gapminder_example <- 
  readxl::read_excel(
    path = "../data/GDPpercapitaconstant2000US.xlsx",
    sheet = "Data")
```

```{r echo=TRUE}
gapminder_example %>% select(1:4) %>% head(n = 4)
```

Task 1: Change that first column's name!

--

Solution

```{r eval=FALSE, echo=TRUE}
gapminder_example %>% 
  rename(country = `Income per person (fixed 2000 US$)`)
```


---

```{r echo=FALSE}
gapminder_example %>% select(1:4) %>% head(n = 4)
```

Task 2: Keep only Germany (hint: there's (West|East) Germany and Germany)

--

Solution

```{r eval=TRUE, echo=TRUE}
gapminder_example %>% 
  rename(country = `Income per person (fixed 2000 US$)`) %>% 
  filter(country %in% 
           c("Germany", "West Germany", "East Germany")) %>% 
  select(1:4)
```


---

```{r echo=FALSE}
gapminder_example %>% select(1:4) %>% head(n = 4)
```

Task 3: Keep only (united, ignore [West|East]) Germany rows and convert to long format

--

Solution

```{r eval=TRUE, echo=TRUE}
gapminder_example %>% 
  rename(country = `Income per person (fixed 2000 US$)`) %>% 
  filter(country == "Germany") %>% 
  gather(2:last_col(), key = year, value = GDP) %>% tail(n = 2)
```

Instead of last_col you could have just counted, too.

---

```{r}
gapminder_example_long <- gapminder_example %>% 
  rename(country = `Income per person (fixed 2000 US$)`) %>% 
  filter(country == "Germany") %>% 
  gather(2:last_col(), key = year, value = GDP)
```

```{r echo=FALSE}
gapminder_example_long %>% head(n = 4)
```

Task 4: Working with the long format, drop all NA-rows

--

Solution

```{r eval=TRUE, echo=TRUE}
gapminder_example_long %>% 
  filter(!is.na(GDP)) %>% head(n = 3)
```

---


```{r echo=FALSE}
gapminder_example_long %>%
  filter(!is.na(GDP)) %>% head(n = 4)
```

Task 5: Compute the average (mean) of Germany's GDP and show yearly deviation.

--

Solution

```{r eval=TRUE, echo=TRUE}
gapminder_example_long %>%
  filter(!is.na(GDP)) %>% mutate(mean_GDP = mean(GDP), 
                                 diff = abs(GDP) - mean_GDP) %>% 
  head(n = 3)
```

---
class: center, middle

<h2> Thanks!</h2>

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
