---
title: "Data Wrangling & Exploration with the Tidyverse in R"
subtitle: "Relational Data"
author: "Johannes Breuer<br />Thomas Ebel<br />Stefan Müller"
date: "2019-05-16"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "../footer-header.css", "../mycss.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include=FALSE}
packages_to_load <- c("knitr",
                      "rmarkdown",
                      "devtools",
                      "tibble",
                      "readr",
                      "kableExtra",
                      "dplyr",
                      "here")

for (i in packages_to_load) {
  if (!require(i, character.only = TRUE)) {
    install.packages(i)
  } else {
    library(i, character.only = TRUE)
  }
}

rm(packages_to_load)

install_github("hadley/emo")
library(emo)

options(htmltools.dir.version = FALSE)

opts_chunk$set(echo = FALSE, fig.align = "center")
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", ", ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$subtitle)`</span></div>
</div>

---

# Combining datasets

In most cases researchers/analysts work with more than one data file.  

The simplest way two combine two tabular datasets is to use `bind_cols()` or `bind_rows()` from `dplyr`. However, to use **`bind_cols()`**, you need to make sure that the datasets contain the **same cases** (rows) in the **same order**. In order to use **`bind_rows()`**, it is necessary to have the **same variables** (with the same names) across the datasets.

`dplyr` offers three categories of so-called *two-table verbs* to combine two tabular datasets:

1. Mutating joins  

2. Filtering joins  

3. Set operations

---

# Superhero data `r ji("bat")``r ji("zap")``r ji("spider_web")``r ji("merman")`

- to illustrate the two-table verbs from `dlpyr` we will use data from the Super Heroes dataset from [*Kaggle*](https://www.kaggle.com/claudiodavi/superhero-set)

```{r superhero data, include = F}
heroes <- read_csv("../../data/other/heroes_information.csv") %>% 
  select(name, Alignment) %>% 
  rename(Name = name) %>%
  distinct() %>% 
  filter(
    Name == "Spider-Man" |
    Name == "Aquaman" |
    Name == "Hellboy"
  )

publishers <- read_csv("../../data/other/heroes_information.csv") %>% 
  select(name, Publisher) %>%
  rename(Name = name) %>%
  distinct() %>% 
  filter(
    Name == "Spider-Man" |
    Name == "Aquaman" |
    Name == "Spawn"
  ) %>% 
  distinct()

# define colors for superheroes
aq_green <- "#008B45"
spi_blue  <- "#0000FF"
spa_black <- "#000000"
he_red <- "#FF0000"

```

.pull-left[
.center[
**Superheroes**
```{r superhero table, echo = FALSE}
heroes %>%
  kable("html")
```
]
]

.pull-right[
.center[
**Publishers**
```{r publishers table, echo = FALSE}
publishers %>%
  kable("html")
```
]
]

---

class: center, middle

# Time for `r ji("coffee")`