---
title: "R Crash Course"
# subtitle: "(according to Hadley Wickham/RStudio)"
author: "Johannes Breuer<br />Thomas Ebel<br />Stefan Müller"
date: "2019-05-15"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
  lib_dir: libs
css: ["default", "robot-fonts", "../footer-header.css", "../mycss.css"]
nature:
  highlightStyle: "github"
highlightLines: true
countIncrementalSlides: false
---
  layout: true

```{r setup, include = FALSE}
packages_to_load <- c("knitr", "rmarkdown", "magrittr", "readr", "ggplot2")

for (i in packages_to_load) {
  if (!require(i, character.only = TRUE)) {
    install.packages(i)
  } else {
    library(i, character.only = TRUE)
  }
}

rm(packages_to_load)

options(htmltools.dir.version = FALSE)

opts_chunk$set(echo = FALSE, fig.align = "center")
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", " & ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$title)`</span></div>
  </div>
  
---
# What is R?
Some people characterize `R` as a programming language, some as a language for doing statistical analysis.

The truth is `R` is used by data analysists and statisticians, but it could also be used, e.g., for application development.

Yet, if you are seeking for one of the most comprehensive tool for data wrangling, analysis, and visualization use `R`

---
# Why should I use R?

- it's powerful
- it's open source / free software
  - great for science
  - great for publicly funded enterprises and authorities
- it can produce nice and clean code
  - this also depends on you
- it has a good integration into other tools
  - RMarkdown
  - LaTeX
  - ...
  
- did I mention, it's free?
  
  
---
# How does R work?
For beginners it is sometimes hard to learn R, but it uses principles and concepts we already know from mathematics... such as,

- vectors
- matrices

Let's be honest: this is why R is so attractive for statistics!

---
# A vector example
```{r vector_example, echo = TRUE}
# create a vector with c()
vector_1 <- c(1, 2, 3, 4, 5)

# print it
vector_1
```

---
# A vector example
```{r vector_example_2, echo = TRUE}
# create a second vector with c()
vector_2 <- c(5, 4, 3, 2, 1)

# merge both vectors
matrix_1 <- cbind(vector_1, vector_2)

# print the resulting matrix
matrix_1
```

--- 
# A matrix example
```{r matrix_example, echo = TRUE}
matrix_1
```

---
# What should I know about R?
First of all, R is a rather flexible tool that can be used in many applications

- without a graphical user interface
  - hey, use it as a calculator on your command line...
- in a graphical user interface
  - provided along the base R installation, e.g., on Windows
  - or more fancy, in an Integrated Development Environment (IDE)
    - in this course, we use RStudio: https://www.rstudio.com/
